A Framework for Substructural Type Systems \thanks{James Wood is supported by an EPSRC Studentship. Robert Atkey is supported by EPSRC grant FIXME.}

### 2022-01-29 tex/esop22/env.tex
[ edit ] squash and address todos

\begin{lemma}[\AgdaFunction{relocate}]\label{thm:env-resize}
  Given an environment $\rho : \grP\gamma \env\V \grQ\delta$ and a $\grPprime$
  and a $\grQprime$ such that $\grPprime \leq \grQprime\plr{\rho.\gr\Psi}$,
  there is also an environment of type $\grPprime\gamma \env\V \grQprime\delta$
  with the same linear map and action on variables.
\end{lemma}
% \begin{proof}
%   The only part of the definition of an environment dependent on $\grP$ or
%   $\grQ$ is the constraint $\grP \leq \grQ\gr\Psi$, which we are able to
%   replace for $\grPprime$ and $\grQprime$.
% \end{proof}
Relocation will be used when pushing environments into subterms in
\secref{sec:traversal}.

@@

When $\V$ supports subusaging, we can construct a $\V$-environment by cases on
the shape of the target context by the following rules, which use the bunched
connectives from \secref{sec:bunched-rules}:
\begin{displaymath}
  \begin{prooftree}[comb,center=false]
    \hypo{I^*}
    \infer1{\alr{} : {} \env\V {\cdot}}
  \end{prooftree}
  \qquad
  \begin{prooftree}[comb,center=false]
    \hypo{\rho : {} \env\V \Delta_l}
    \hypo{\sep}
    \hypo{\sigma : {} \env\V \Delta_r}
    \infer3{\alr{\rho,\sigma} : {} \env\V \Delta_l, \Delta_r}
  \end{prooftree}
  \qquad
  \begin{prooftree}[comb,center=false]
    \hypo{\gr r\cdot\plr{M : {} \sdtstile{}\V A}}
    \infer1{\alr{M} : {} \env\V \gr rA}
  \end{prooftree}
\end{displaymath}
Left to right, we can create an environment into the empty context
when all usage annotations on the source context are $\gr0$; we can
create an environment into a concatenated context when we can
additively split up the annotations of the source context and produce
environments into both halves from the split sources; and we can
create an environment into a singleton context $\gr rA$ when we can divide the
source context by $\gr r$ and produce a $\V$-value in the divided context of the
appropriate type.

@@

We could have used these rules to inductively define what environments are.
However, we found that this was difficult to work with. It is often easier to do
linear algebraic proofs separately from the rest of an environment.
For example, for identity and composition of environments (below), \defref{def:lr-env}
is easier to use because we can rely on the identity and composition of linear
maps.
Concretely, an inductive proof of identity would, for example, involve
constructing an environment of type
$\grP\gamma, \grQ\delta \env\V \grP\gamma, \grQ\delta$ by constructing
environments of types $\grP\gamma, \gr0\delta \env\V \grP\gamma$ and
$\gr0\gamma, \grQ\delta \env\V \grQ\delta$.
These are not identity environments, so we would have to strengthen the
induction hypothesis.

@@

\paragraph{Identity and Composition} $\V$-environments map variables
to $\V$-things, so they do not {\it a priori} support identities and
composition. However, by assuming extra structure we can obtain
identities and composition for environments:
\bob{Need to mention when these are real id/comp}

\begin{lemma}[Identity environment]\label{thm:env-id}
  Given a function $\mathrm{vr} : \forallb{{\sqni} \dotto \V}$, for any
  $\Gamma$ we have an environment of type $\Gamma \env\V \Gamma$.
\end{lemma}

\begin{lemma}[Composition of environments]\label{thm:env-comp}
  Given a function
  \begin{displaymath}
    \mathrm{lift} : \plr{\rho : \grP\gamma \env\U \grQ\delta} \to \forallb{\V\,\grQ\delta \dotto \W\,\grP\gamma}
  \end{displaymath}
  then we can compose environments $\rho : \Gamma \env\U \Delta$ and
  $\rho' : \Delta \env\V \Theta$ into an environment
  $\rho; \rho' : \Gamma \env\W \Theta$.
\end{lemma}

@@

\bob{Forward reference to where we use identity and composition}

@@

\paragraph{Renameability}
Renamings, i.e. $\sqni$-environments, are a particularly important case of
environments.
As in the work of \citet{FPT99}, presheaves over renamings are an important
class of open families.
In Agda code, we abbreviate $\env\sqni$ (which would usually be
\AgdaOperator{\AgdaRecord{[}}\AgdaSpace{}%
\AgdaOperator{\AgdaRecord{\AgdaUnderscore{}$\sqni$\AgdaUnderscore{}}}\AgdaSpace{}%
\AgdaOperator{\AgdaRecord{]\AgdaUnderscore{}$\Rightarrow^e$\AgdaUnderscore{}}}%
) as \AgdaFunction{\AgdaUnderscore{}$\Rightarrow^r$\AgdaUnderscore{}}.

\todo{Sort this out: we get a category of renamings, and the comonad a la AACMM}

@@

A renameable notion of value gives rise to a renameable notion of environment,
essentially by renaming each contained value in an appropriate way.
On the other side, environments are always renameable in their codomains.

\begin{lemma}[\AgdaFunction{ren\textasciicircum{}Env}]\label{thm:env-ren}
  If $\plr{-} \sdtstile{}\V A$ is renameable for all $A$, then so is
  $\plr{-} \env\V \Delta$ for all $\Delta$.
\end{lemma}
\bob{Proof probably isn't needed}
\begin{proof}
  For renameable $\V$, we can implement $\mathrm{lift}$ as in
  \cref{thm:env-comp} with $\U \coloneqq {\sqni}$ and
  $\V \coloneqq \W \coloneqq \V$.
  Renaming the environment then becomes an instance of environment composition.
\end{proof}
\begin{lemma}\label{thm:env-postren}
  $\Gamma \env\V \plr{-}$ is renameable for all $\V$ and $\Gamma$.
\end{lemma}
\begin{proof}
  Corollary of \cref{thm:env-comp}.
  \todo{Or we may be removing environment composition.}
\end{proof}

>>>>

\begin{lemma}[\AgdaFunction{relocate}]\label{thm:env-resize}
  Given an environment $\rho : \grP\gamma \env\V \grQ\delta$ and a $\grPprime$
  and a $\grQprime$ such that $\grPprime \leq \grQprime\plr{\rho.\gr\Psi}$,
  there is also an environment of type $\grPprime\gamma \env\V \grQprime\delta$
  with the same linear map and action on variables.
\end{lemma}
% \begin{proof}
%   The only part of the definition of an environment dependent on $\grP$ or
%   $\grQ$ is the constraint $\grP \leq \grQ\gr\Psi$, which we are able to
%   replace for $\grPprime$ and $\grQprime$.
% \end{proof}
Relocation will be used when pushing environments into subterms in
\cref{sec:traversal}.

@@

When $\V$ supports subusaging, we can construct a $\V$-environment by cases on
the shape of the target context by the following rules, which use the bunched
connectives from \cref{sec:bunched-rules}:
\begin{displaymath}
  \begin{prooftree}[comb,center=false]
    \hypo{I^*}
    \infer1{\alr{} : {} \env\V {\cdot}}
  \end{prooftree}
  \qquad
  \begin{prooftree}[comb,center=false]
    \hypo{\rho : {} \env\V \Delta_l}
    \hypo{\sep}
    \hypo{\sigma : {} \env\V \Delta_r}
    \infer3{\alr{\rho,\sigma} : {} \env\V \Delta_l, \Delta_r}
  \end{prooftree}
  \qquad
  \begin{prooftree}[comb,center=false]
    \hypo{\gr r\cdot\plr{M : {} \sdtstile{}\V A}}
    \infer1{\alr{M} : {} \env\V \gr rA}
  \end{prooftree}
\end{displaymath}
Left to right, we can create an environment into the empty context
when all usage annotations on the source context are $\gr0$; we can
create an environment into a concatenated context when we can
additively split up the annotations of the source context and produce
environments into both halves from the split sources; and we can
create an environment into a singleton context $\gr rA$ when we can divide the
source context by $\gr r$ and produce a $\V$-value in the divided context of the
appropriate type.

@@

We could have used these rules to inductively define what environments are.
However, we found that this was difficult to work with. It is often easier to do
linear algebraic proofs separately from the rest of an environment.
For example, for identity and composition of environments (below), \cref{def:lr-env}
is easier to use because we can rely on the identity and composition of linear
maps.
Concretely, an inductive proof of identity would, for example, involve
constructing an environment of type
$\grP\gamma, \grQ\delta \env\V \grP\gamma, \grQ\delta$ by constructing
environments of types $\grP\gamma, \gr0\delta \env\V \grP\gamma$ and
$\gr0\gamma, \grQ\delta \env\V \grQ\delta$.
These are not identity environments, so we would have to strengthen the
induction hypothesis.

@@

%\paragraph{Identity and Composition} $\V$-environments map variables
%to $\V$-things, so they do not {\it a priori} support identities and
%composition. However, by assuming extra structure we can obtain
%identities and composition for environments:
%\bob{Need to mention when these are real id/comp}
%
%\begin{lemma}[Identity environment]\label{thm:env-id}
%  Given a function $\mathrm{vr} : \forallb{{\sqni} \dotto \V}$, for any
%  $\Gamma$ we have an environment of type $\Gamma \env\V \Gamma$.
%\end{lemma}
%
%\begin{lemma}[Composition of environments]\label{thm:env-comp}
%  Given a function
%  \begin{displaymath}
%    \mathrm{lift} : \plr{\rho : \grP\gamma \env\U \grQ\delta} \to \forallb{\V\,\grQ\delta \dotto \W\,\grP\gamma}
%  \end{displaymath}
%  then we can compose environments $\rho : \Gamma \env\U \Delta$ and
%  $\rho' : \Delta \env\V \Theta$ into an environment
%  $\rho; \rho' : \Gamma \env\W \Theta$.
%\end{lemma}

@@

%\bob{Forward reference to where we use identity and composition}

@@

\paragraph{Renameability}
Renamings, i.e. $\sqni$-environments, are a particularly important case of
environments.
Renamings form a category, with identity and composition following from the
identity and composition of linear maps.
As in the work of \citet{FPT99}, presheaves over renamings are an important
class of open families.
In Agda code, we abbreviate $\env\sqni$ (which would usually be
\AgdaOperator{\AgdaRecord{[}}\AgdaSpace{}%
\AgdaOperator{\AgdaRecord{\AgdaUnderscore{}$\sqni$\AgdaUnderscore{}}}\AgdaSpace{}%
\AgdaOperator{\AgdaRecord{]\AgdaUnderscore{}$\Rightarrow^e$\AgdaUnderscore{}}}%
) as \AgdaFunction{\AgdaUnderscore{}$\Rightarrow^r$\AgdaUnderscore{}}.

@@

A renameable notion of value gives rise to a renameable notion of environment,
essentially by renaming each contained value in an appropriate way.
On the other side, environments admit renamings of their codomains in the
opposite direction to that given by renameability.

\begin{lemma}[\AgdaFunction{ren\textasciicircum{}Env}]\label{thm:env-ren}
  If $\plr{-} \sdtstile{}\V A$ is renameable for all $A$, then so is
  $\plr{-} \env\V \Delta$ for all $\Delta$.
\end{lemma}
%\bob{Proof probably isn't needed}
%\begin{proof}
%  For renameable $\V$, we can implement $\mathrm{lift}$ as in
%  \cref{thm:env-comp} with $\U \coloneqq {\sqni}$ and
%  $\V \coloneqq \W \coloneqq \V$.
%  Renaming the environment then becomes an instance of environment composition.
%\end{proof}
\begin{lemma}\label{thm:env-postren}
  From $\Gamma \env\V \Delta$ and $\Delta \env{\sqni} \Theta$, we get
  $\Gamma \env\V \Theta$.
\end{lemma}
\begin{proofsketch}
  Notice that the lookup component of an environment maps variables in the
  codomain to values in the domain.
  We can apply the renaming to these variables.
\end{proofsketch}
%\begin{proof}
%  Corollary of \cref{thm:env-comp}.
%  \todo{Or we may be removing environment composition.}
%\end{proof}

### 2022-01-29 tex/esop22/examples.tex
[ edit ] squash and address todos

TODO: introductory paragraph

\bob{Reification dumped here for now}
When \AgdaBound{$\V$} supports identity environments (as per \cref{thm:env-id}),
we can simplify our obligations in giving a \AgdaRecord{Semanitics} by coercing
\AgdaFunction{Kripke} functions into just
\AgdaBound{$\C$}-values in an extended context.

@@

\begin{lemma}[\AgdaFunction{reify}]\label{thm:reify}
  If $\V$ is an open family such that there is a function
  $v : \forallb{{\sqni} \dotto \V}$, we get a function of type
  $\forallb{\mathrm{Kripke}\,\V\,\C \dotto \mathrm{Scope}\,\C}$ for any $\C$.
\end{lemma}
\begin{proof}
  Let $b : \mathrm{Kripke}\,\V\,\C\,\Delta\,\Gamma\,A$.
  That is, $b$ is a Kripke function yielding $\C$-computations
  We want to apply $b$ so as to get a $\C\,\plr{\Gamma, \Delta}\,A$.
  Let $\grP\gamma = \Gamma$ and $\grQ\delta = \Delta$.
  The $\Box^r$ in the type of $b$ allows us to reverse-rename $\Gamma$ to
  $\Gamma, \gr0\delta$.
  Then we give the $\wand$-function an argument in context
  $\gr0\gamma, \Delta$, noting that
  $\plr{\Gamma, \gr0\delta} + \plr{\gr0\gamma, \Delta} = \plr{\Gamma, \Delta}$,
  as we wanted from the result.
  The argument needs type $\gr0\gamma, \Delta \env\V \Delta$.
  We produce this through the composition
  $\gr0\gamma, \Delta \env\V \gr0\gamma, \Delta \env{\sqni} \Delta$,
  using the instance of composition mentioned in \cref{thm:env-postren}.
  The $\V$-environment follows from $v$ and \cref{thm:env-id}, while the
  renaming is the complement to that used on $\Box^r$.
\end{proof}

@@

We will take semantics of this system into
\emph{world-indexed relations}~\cite{AbelBernardy2020,context-constrained}.
A world-indexed relation over a poset of worlds \AgdaBound{W} is a set over
which
we have a \AgdaBound{W}-indexed binary relation satisfying a presheaf-like
property with respect to the order on \AgdaBound{W}.

\ExecuteMetaData[\WReltex]{WRel}

@@

Morphisms between world-indexed relations \AgdaBound{R} and \AgdaBound{S}
consist of a mapping between the underlying sets such that that mapping
preserves relatedness from \AgdaBound{R} to \AgdaBound{S}.

\ExecuteMetaData[\WReltex]{WRelMor}

\todo{Define big intersection.}
When the poset of worlds forms a (relational) commutative monoid, such
world-indexed relations support a symmetric monoidal closed structure, with
objects denoted \AgdaFunction{I$^R$},
\AgdaFunction{\AgdaUnderscore{}$\otimes^R$\AgdaUnderscore{}}, and
\AgdaFunction{\AgdaUnderscore{}$\multimap^R$\AgdaUnderscore{}},.
These reuse the bunched connectives \AgdaRecord{$I^*$}, \AgdaRecord{$\sep$}, and
\AgdaRecord{$\wand$}, now over worlds rather than contexts.

%\ExecuteMetaData[\WReltex]{IR}
%\ExecuteMetaData[\WReltex]{tensorR}
%\ExecuteMetaData[\WReltex]{lollyR}

The final piece of sematics we need is a \emph{bang} operator.
We allow the
semantic \emph{bang} to be an arbitrary annotation-indexed functor on
world-indexed relations.
This functor must respect all of the structure on the indices, making it a
graded comonad over multiplication, as well as being lax monoidal at any
particular index \AgdaBound{r}.

\ExecuteMetaData[\WReltex]{Bang}

@@

To associate semantics to syntax, we start as standard by associating
world-indexed relations to types.
We also extend the semantics of types to contexts, using \AgdaFunction{I$^R$},
\AgdaFunction{$\otimes^R$}, and \AgdaField{!$^R$} to interpret the empty
context, concatenation, and usage annotations on singletons, respectively.

\ExecuteMetaData[\WReltex]{sem}

The semantics of a term is then to be a morphism from the interpretation of the
context to the interpretation of the term's type.

\ExecuteMetaData[\WReltex]{sem-vdash}

Variables are given semantics by \AgdaFunction{lookup$^R$} (definition omitted).

@@

Now, we give a \AgdaRecord{Semantics}.
The choice of \AgdaBound{$\V$} as
\AgdaRecord{\AgdaUnderscore{}$\sqni$\AgdaUnderscore{}} is somewhat arbitrary,
given that a standard denotational semantics would not use intermediate
environments in the same sense as renaming and substitution, but allows us to
reuse the standard facts that variables support renaming and identity
environments.
With this choice of \AgdaBound{$\V$} and \AgdaBound{$\C$}, we interpret
environment entries by \AgdaFunction{lookup$^R$}.
Meanwhile, for the logical rules, we ignore environments by using
\AgdaFunction{reify} to just deal with morphisms in an extended context.
As such, $\lambda$-abstractions are easy to interpret, while applications
require some massaging to remove the extension by an empty context, followed by
some plumbing to split the interpretation of the context according to the usage
constraints and feed the interpretation of the argument \AgdaBound{n$'$} into
the interpretation of the function \AgdaBound{m$'$}.

@@

In order to map open terms to interpretations, we take the action of the
semantics and give the identity renaming as the starting environment.

\ExecuteMetaData[\WReltex]{wrel}

>>>>

In this section, we provide three example uses of semantic traversals: generic
renaming and substitution, a usage elaborator, and a denotational semantics.
The reader is also encouraged to see the far greater range of examples in the
work of \citet{AACMM21}, which should adapt to our usage-annotated setting.
Renaming and substitution are essential results, while the latter two examples
focus on usage annotations.

A result we will use throughout this section is \emph{reification}.
When we have an index-preserving mapping from usage-checked variables to
$\V$-environments, we can construct environments of the form
$\Gamma \env\V \Gamma$ (identity environments) for all $\Gamma$.
This lets us write the \AgdaFunction{reify} function, which  simplifies our
obligations in giving a \AgdaRecord{Semanitics} by coercing
\AgdaFunction{Kripke} functions into just
\AgdaBound{$\C$}-values in an extended context.

@@

\begin{lemma}[\AgdaFunction{reify}]\label{thm:reify}
  If $\V$ is an open family such that there is a function
  $v : \forallb{{\sqni} \dotto \V}$, we get a function of type
  $\forallb{\mathrm{Kripke}\,\V\,\C \dotto \mathrm{Scope}\,\C}$ for any $\C$.
\end{lemma}
\begin{proof}
  Let $b : \mathrm{Kripke}\,\V\,\C\,\Delta\,\Gamma\,A$.
  That is, $b$ is a Kripke function yielding $\C$-computations
  We want to apply $b$ so as to get a $\C\,\plr{\Gamma, \Delta}\,A$.
  Let $\grP\gamma = \Gamma$ and $\grQ\delta = \Delta$.
  The $\Box^r$ in the type of $b$ allows us to reverse-rename $\Gamma$ to
  $\Gamma, \gr0\delta$.
  Then we give the $\wand$-function an argument in context
  $\gr0\gamma, \Delta$, noting that
  $\plr{\Gamma, \gr0\delta} + \plr{\gr0\gamma, \Delta} = \plr{\Gamma, \Delta}$,
  as we wanted from the result.
  The argument needs type $\gr0\gamma, \Delta \env\V \Delta$.
  We produce this via \cref{thm:env-postren} from an environment
  $\rho : \gr0\gamma, \Delta \env\V \gr0\gamma, \Delta$ created using $v$
  and a renaming which is the complement to that used on $\Box^r$.
\end{proof}

@@

We will take semantics of this system into
\emph{world-indexed relations}~\cite{AbelBernardy2020,context-constrained}.
A world-indexed relation (\AgdaRecord{WRel}) over a poset of worlds
\AgdaBound{W} is a set over which
we have a \AgdaBound{W}-indexed binary relation satisfying a presheaf-like
property with respect to the order on \AgdaBound{W}.
The rendered Agda code for world-indexed relations and constructions on them
can be found in \cref{sec:den-sem-extras}.

@@

Morphisms (\AgdaRecord{WRelMor}) between world-indexed relations \AgdaBound{R}
and \AgdaBound{S} consist of a mapping between the underlying sets such that, at
each fixed world \AgdaBound{w}, the mapping preserves relatedness from
\AgdaBound{R} to \AgdaBound{S}.

When the poset of worlds forms a (relational) commutative monoid, such
world-indexed relations support a symmetric monoidal closed structure, with
objects denoted \AgdaFunction{I$^R$},
\AgdaFunction{\AgdaUnderscore{}$\otimes^R$\AgdaUnderscore{}}, and
\AgdaFunction{\AgdaUnderscore{}$\multimap^R$\AgdaUnderscore{}},.
These reuse the bunched connectives \AgdaRecord{$I^*$}, \AgdaRecord{$\sep$}, and
\AgdaRecord{$\wand$}, now over worlds rather than contexts.

The final piece of sematics we need is a \emph{bang} operator.
We allow the
semantic \emph{bang} to be an arbitrary annotation-indexed functor on
world-indexed relations.
This functor must respect all of the structure on the indices, making it a
graded comonad over multiplication, as well as being lax monoidal at any
particular index \AgdaBound{r}.
These laws are listed in \cref{sec:den-sem-extras}.

@@

%To associate semantics to syntax, we start as standard by associating
%world-indexed relations to types.
%We also extend the semantics of types to contexts, using \AgdaFunction{I$^R$},
%\AgdaFunction{$\otimes^R$}, and \AgdaField{!$^R$} to interpret the empty
%context, concatenation, and usage annotations on singletons, respectively.
%
%\ExecuteMetaData[\WReltex]{sem}
%
%The semantics of a term is then to be a morphism from the interpretation of the
%context to the interpretation of the term's type.
%
%\ExecuteMetaData[\WReltex]{sem-vdash}

The semantics of a type is given by \AgdaFunction{$\llbracket$\_$\rrbracket$},
which maps into world-indexed relations.
The function type is interpreted using \AgdaFunction{$\multimap^R$} and
\AgdaField{!$^R$}.
Contexts are interpreted by \AgdaFunction{$\llbracket$\_$\rrbracket^c$}, using
\AgdaFunction{$\otimes^R$} and \AgdaFunction{I$^R$}.
Terms are interpreted as morphisms by the open family
\AgdaFunction{$\llbracket$\_$\vdash$\_$\rrbracket$}.
Variables are interpreted by \AgdaFunction{lookup$^R$} (definition omitted).

@@

Now we give a \AgdaRecord{Semantics}.
The choice of \AgdaBound{$\V$} as
\AgdaRecord{\AgdaUnderscore{}$\sqni$\AgdaUnderscore{}} is somewhat arbitrary,
given that a standard denotational semantics would not use intermediate
environments in the same sense as renaming and substitution, but it allows us to
reuse the standard facts that variables support renaming and identity
environments.
With this choice of \AgdaBound{$\V$} and \AgdaBound{$\C$}, we interpret
environment entries by \AgdaFunction{lookup$^R$}.
Meanwhile, for the logical rules, we ignore environments by using
\AgdaFunction{reify} to just deal with morphisms in an extended context.
As such, $\lambda$-abstractions are easy to interpret, while applications
require some massaging to remove the extension by an empty context, followed by
some plumbing to split the interpretation of the context according to the usage
constraints and feed the interpretation of the argument \AgdaBound{n$'$} into
the interpretation of the function \AgdaBound{m$'$}.

@@

Then, the semantics of terms is given by the function
\AgdaFunction{semantics}\AgdaSpace{}\AgdaFunction{Wrel}\AgdaSpace{}%
\AgdaFunction{1$^r$}, where \AgdaFunction{1$^r$} is the identity renaming.
%In order to map open terms to interpretations, we take the action of the
%semantics and give the identity renaming as the starting environment.
%
%\ExecuteMetaData[\WReltex]{wrel}

### 2022-01-29 tex/esop22/semantics.tex
[ edit ] squash and address todos

% In \secref{sec:terms}, we defined \AgdaFunction{Scope} to let a
% judgement-indexed family admit context extensions. However, a key
% component of our generic semantic traversal is to make use of the open
% family \AgdaBound{$\V$} of \emph{values}, which are the sort of thing
% we store in an environment.  The definition \AgdaFunction{Kripke}
% gives an alternative to \AgdaFunction{Scope} which interprets the
% newly bound variables via a requirement of $\V$-values rather than
% extra assumptions for the $\C$-computation.

@@

\AgdaFunction{Kripke} is functorial in the $\C$ argument, as witnessed
by the \AgdaFunction{mapKC} function, which is essentially
post-composition:

>>>>

% In \cref{sec:terms}, we defined \AgdaFunction{Scope} to let a
% judgement-indexed family admit context extensions. However, a key
% component of our generic semantic traversal is to make use of the open
% family \AgdaBound{$\V$} of \emph{values}, which are the sort of thing
% we store in an environment.  The definition \AgdaFunction{Kripke}
% gives an alternative to \AgdaFunction{Scope} which interprets the
% newly bound variables via a requirement of $\V$-values rather than
% extra assumptions for the $\C$-computation.

@@

\AgdaFunction{Kripke} is functorial in the \AgdaBound{$\C$} argument,
as witnessed by the \AgdaFunction{mapKC} function, which is essentially
post-composition:

### 2022-01-29 tex/esop22/syntax.tex
[ edit ] squash and address todos

We take the insights of the previous section and use them to build a
generic framework for posemiring-annotated substructural systems in
Agda. We will first show \emph{descriptions} of systems, which are
comprised of rules that have premises combined using the bunched
combinators. We then show how to construct the Agda data type of
intrinsically well scoped, typed, and resourced terms for any given
system of our framework. We use the prototypical system from
\figref{fig:lr-comb} as a running example. \secref{sec:other-syntaxes}
presents further examples that our framework can express.

@@

% \paragraph{\figref{fig:lr-comb} as a \AgdaRecord{System}.}

As an example, we transcribe the system defined in
\figref{fig:lr-comb} into a description.  We give the set of types of
this system as a data type \AgdaDatatype{Ty} (together with a base
type \AgdaInductiveConstructor{$\iota$}). We assume that there is a
posemiring \AgdaInductiveConstructor{Ann} in scope for the
annotations.There is one label for each instantiation of a logical
rule, but the labels contain no further information about subterms or
restrictions on the context. This will be provided when we associate
labels with \AgdaRecord{Rule}s in a \AgdaRecord{System}.

@@

Compared to \figref{fig:lr-comb}, modulo the Agda notation, we can see
that the fundamental structure has been preserved: the rules match
one-to-one, and the bunched premises are the same. A major difference
is that we do not include a counterpart to the
\AgdaInductiveConstructor{var} rule in a
\AgdaRecord{System}. Variables are common to all the systems
representable in our framework.

\subsection{Terms of a \AgdaRecord{System}.}\label{sec:terms}

@@

Terms defined like this are difficult to write by hand, due to the
need for proofs that the usage contexts are handled correctly for the rules used.
We will see a way of automating these proofs in \cref{sec:usage-elaborator}.

>>>>

We take the insights of the previous section and use them to build a
generic framework for posemiring-annotated substructural systems in
Agda. We will first show \emph{descriptions} of systems, which are
comprised of rules that have premises combined using the bunched
combinators. We then show how to construct the Agda data type of
intrinsically well scoped, typed, and resourced terms for any given
system of our framework. We use the prototypical system from
\cref{fig:lr-comb} as a running example. \Cref{sec:other-syntaxes}
presents further examples that our framework can express.

@@

% \paragraph{\Cref{fig:lr-comb} as a \AgdaRecord{System}.}

As an example, we transcribe the system defined in
\cref{fig:lr-comb} into a description.  We give the set of types of
this system as a data type \AgdaDatatype{Ty} (together with a base
type \AgdaInductiveConstructor{$\iota$}). We assume that there is a
posemiring \AgdaInductiveConstructor{Ann} in scope for the
annotations.There is one label for each instantiation of a logical
rule, but the labels contain no further information about subterms or
restrictions on the context. This will be provided when we associate
labels with \AgdaRecord{Rule}s in a \AgdaRecord{System}.

@@

Compared to \cref{fig:lr-comb}, modulo the Agda notation, we can see
that the fundamental structure has been preserved: the rules match
one-to-one, and the bunched premises are the same. A major difference
is that we do not include a counterpart to the
\AgdaInductiveConstructor{var} rule in a
\AgdaRecord{System}. Variables are common to all the systems
representable in our framework.

\subsection{Terms of a \AgdaRecord{System}}\label{sec:terms}

@@

Terms defined like this are difficult to write by hand, due to the
need for proofs that the usage contexts are handled correctly for the rules used.
We give an example of a hand-written term in \cref{sec:explicit-term}, and will
see a way of automating the necessary proofs in \cref{sec:usage-elaborator}.
